Accelerated convergence of convective simulations using boundary value problems

### 2018-01-12 tex/bvp_paper.tex
basically completes rough draft of paper, need to get new data in and see how it looks

\begin{abstract}
We present a method for using coupling Boundary value problems (BVPs) with Initial value problems (IVPs)
in order to achieve thermally converged convective solutions on dynamical timescales, rather than the
long thermal timescale.  We demonstrate the similarity between the solution reached via BVP and the
solution reached by a long thermal rundown of the IVP, and demonstrate that this method works at a
large range of supercriticalities.  We use this method to achieve converged solutions at high Ra,
and discuss its extension to more complex scenarios, such as stratified, compressible convection.
\end{abstract}
\maketitle

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales. Granules on the
solar surface overturn on the order of 10 minutes, whereas deep motions in the Sun are likely at
low Mach number and constrained by the solar rotation rate of $\sim$1 month.  
Both of these dynamical timescales are vastly shorter than the Sun's average timescale of energy transport,
which is the Kelvin-Helmholtz timescale of nearly $3 \times 10^7$ \emph{years} \cite{stix2003}. 
As simulations aim to model natural convection
by increasing into the high-Rayleigh Number (Ra) regime, where diffusive timescales are much
longer than dynamical timescales \cite{Anders&Brown2017}, achieving converged simulations will 
require runs which span a greater number of convective timescales in order to thermally converge.
Furthermore, with increasing Ra and decreasing diffusivities, motions become more turbulent
and require finer grid meshes and shorter timesteps to resolve turbulent motions, increasing the simulation
time required for each overturn timescale.  
These two effects conspire to make achieving thermally converged, high-Ra, astrophysically interesting
simulations an intractable problem using modern numerical tools.

Prior studies of stratified convection in which a convective layer lies between stable layers
have used the knowledge of Mixing Length Theory (MLT) to adjust the initial thermodynamic structure of
the atmosphere to a state which is closer to the adiabat chosen by convection \cite{brandenburg&all2005}.
However, most studies of convection do not contain stable layers above and below the convection
zone, and the presence of hard boundaries and the boundary layers that they form make it difficult to
know the correct evolved adiabat \emph{a priori}.

Here we present a method for using simple boundary value problems (BVPs), 
along with information about the evolved flow fields,
to fast-forward the slow thermal evolution of convecting simulations.  
We run two sets of experiments: one in which we allow convective simulations to evolve for a
full thermal timescale before taking measurements, and another in which we employ a fast-forwarding,
BVP technique which occurs on dynamical timescales. We compare these two sets of simulations to
show the validity of the BVP technique.  Then, we use the BVP technique to run simulations
at high Ra, in the regime where running for thermal timescales becomes computationally intractable.

\section{Experiment}
\label{sec:experiment}
We adopt the Oberbeck-Boussinesq approximation.  Under this choice, the
fluid has constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The variables of the fluid that are evolved are the velocity,
$\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$, the temperature $T = T_0 + T_1$, and the pressure.
The density of the fluid is a constant $\rho_0$, except on the
term where the constant gravitational acceleration, $\bm{g} = - g\hat{z}$ acts in the vertical momentum equation, 
in which case it is $\rho = \rho_0(1  - \alpha T_1)$.  
Under these choices, the equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:dim_incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
We non-dimensionalize these equations such that length scales are in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and time is in units of the freefall timescale ($L_z / v_{\text{ff}}$), where the freefall velocity is
$v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$.
We further re-arrange the momentum equation by introducing a reduced pressure
$\varpi \equiv P / \rho_0 + \phi + |\bm{u}|^2 / 2$, where $\bm{g} = -\grad \phi$ is the gravitational potential.
Despite containing a nonlinear $\bm{u}$ component, the nature of $P$ in \RB convection, as a (whatever it's called)
allows us to treat $\varpi$ as a linear unknown, without a need to resolve its different components.
We time evolve the equations in the form,
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega}
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity. We find this form of the momentum equation to
be slightly faster numerically than the standard form in Eq. (\ref{eqn:dim_bouss_momentum}).
The dimensionless control parameters are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\chi} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\chi}, \qquad \text{Pr} = \frac{\nu}{\chi}.
\end{equation}
The dimensionless vertical extent of the domain is $z = [-1/2, 1/2]$, and at the boundaries
we impose no-slip, impenetrable boundary conditions such that $w = u = v = 0$ at $z = \pm 1/2$.
At the lower boundary, we employ a fixed flux condition such that $\partial T_1 / \partial z = 0$
at $z = -1/2$, and we impose a fixed temperature condition $T_1 = 0$ at $z = 1/2$. Both
horizontal directions are periodic, extending over a range $x, y = [0, \Gamma]$, where
the aspect ratio is $\Gamma = 2$.

The chosen thermal boundary conditions at the upper and lower plates
determine key quantities of the evolved state.
Studies of incompressible, Boussinesq, \RB convection often
employ fixed temperature (Dirichlet) or fixed heat flux
(Neumann) boundary conditions at both plates.  
Dirichlet conditions represent plates of infinite conductivity,
whereas Neumann conditions model plates of finite conductivity.  
In both cases, choosing symmetric boundary conditions maintains overall system symmetry, 
and despite evolving towards different thermal structures, both types of conditions
transport heat in the same manner \cite{johnston&doering2009}.
Studies of convection which aim to model
astrophysical systems such as stars often employ a mixture of these
two types of boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017}.  
The flux at the lower boundary is fixed, modeling
the constant energy generation of the stellar core, 
while the outer boundary condition is held at a fixed temperature,
modeling the surface of a star which must output the energy generated internally.
This setup is a useful model for understanding natural
systems, but simulations which employ these boundary conditions suffer from a long 
thermal relaxation as the atmosphere loses energy and approaches the adiabat chosen by the
Dirichlet condition.  We choose these conditions in part to better understand them, and in
part because these conditions minimize the number of assumptions that must be made in
setting up the boundary value problem.

@@

It is a reasonable assumption to naively guess that convection will evolve the mean temperature profile that
drives it over the course of a thermal diffusion time, $t_\chi \approx \mathcal{P}^{-1}$. Thus, as Ra
increases, the timescale for achieving a thermally converged mean temperature profile becomes intractable.
This prohibitively long thermal timescale in a Direct Numerical Simulation (DNS) 
can be skipped by coupling the DNS with a simple Boundary Value Problem
(BVP) solve. By using information about the evolved dynamical state of the atmosphere,
a BVP can be used to solve for the evolved atmospheric state on short timescales.
A comparison of a simulation running for the thermal timescale, and another where a BVP is used
to fast-forward atmospheric evolution, are shown in Fig \ref{fig:time_trace}a\&b.

The Boussinesq BVP in essence contains equations of hydrostatic balance and thermal equilibrium,
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi} - \angles{T_1}\hat{z} = \angles{\bm{u}\times\bm{\omega}},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\angles{wT_1} - \frac{1}{\text{Pr}\text{Ra}}\frac{\partial^2}{\partial z^2} \angles{T_1} = 0,
	\label{eqn:bouss_BVP_energy}
\end{gather}
where $\angles{A}$ represents a time- and horizontally averaged profile of the quantity $A$.  
These
equations arise from assuming that the atmosphere is in a steady state ($\partial_t \rightarrow 0$),
then taking time and horizontal averages of Eqns (\ref{eqn:bouss_momentum}\&\ref{eqn:bouss_energy}) and
neglecting terms that vanish due to the symmetry of the problem.
Convective flows
are perturbations around a thermal profile defined by these equations in the proper evolved, statistically stationary state.

@@

By definition, the profile of $F_{conv}$ is not in its time stationary early in the DNS.  
In fact, as the atmosphere approaches the
isotherm specified by the upper (fixed $T$) boundary condition, the motions display an asymmetric flux as energy
leaks through the upper boundary condition (Fig. \ref{fig:time_trace}c) in order to reach a lower temperature state.
In order to construct the evolved convective flux from the current fluxes in the atmosphere,
we assume that the atmospheric dynamics have properly developed the thickness of the boundary layers, 
or that the quantities
\begin{equation}
f_{\text{conv}} = \frac{F_{\text{conv}}}{F_{\text{tot}}}\qquad
f_{\text{cond}} = \frac{F_{\text{cond}}}{F_{\text{tot}}}
\end{equation}
early in the evolution are the same as those in the final steady state solution. By definition, the
evolved total flux through the atmosphere must be the same as the flux entering the atmosphere at the
bottom boundary, $F_{\text{tot, steady}} = F_{\text{bot}}$.  The proper enthalpy flux profile to feed in
to Eqn. (\ref{eqn:bouss_BVP_energy}) is then 
$\angles{w T_1}_{\text{steady}} = F_{\text{tot, steady}} \cdot f_{\text{conv}}$.  Presumably, over the
course of the simulation, the velocity field and temperature fluctuations naturally scale like
$d(z) = \angles{w T_1}_{\text{steady}}/\angles{w T_1}$, and so it is necessary to multiply the velocity
field and the fluctuations around the mean in $T_1$ by $\sqrt{d}$.  This also means that
$\angles{\bm{u}\times\bm{\omega}}_{\text{steady}} = d \angles{\bm{u}\times\bm{\omega}}$. Once these profiles
are appropriately adjusted, they can be used in the BVP solve to find the mean profiles of $T$ and $\varpi$.

In general, the BVP solve is completed in the following steps:
\begin{enumerate}
\item Run the convective DNS. Once the convection achieves a volume-averaged Re of $\sqrt{\text{Ra}/\text{Ra}_{\text{crit}}}$,
wait for 50 time units.  Then, start taking the averages of $F_{\text{conv}}$ and $F_{\text{tot}}$, waiting either
30 time units or until the profiles are converged to 1 part in 1000, whichever is a more difficult. This ensures that
the profiles being used in the BVP are steady and smooth, and that they sample the full behavior of the convection.
\item Construct $\angles{w T_1}_{\text{steady}}$ and $\angles{\bm{u}\times\bm{\omega}}_{\text{steady}}$
from the flux profiles.
\item Solve for $\angles{T_1}$ and \angles{\varpi} of the
evolved state.  Adjust the mean profiles in the IVP.
\item Multiply the velocity field and the fluctuations in $T_1$ about its horizontal average by $\sqrt{d}$ in the IVP. 
\item Continue running the IVP for 50 time units to allow for the velocities to equilibrate to their new background state.
\end{enumerate}

\subsection{Numerics}
We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to time-evolve  
(\ref{eqn:incompressible})-(\ref{eqn:bouss_energy}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
The temperature field is decomposed as $T = T_0(z) + T_1(x, y, z, t)$
and the velocity is $\bm{u} = w\bm{\hat{z}} + u\bm{\hat{x}} + v\bm{\hat{y}}$.
In our 2D runs, $v = 0$.
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Domain sizes range from
32x128 coefficients at the lowest values of 
Ra to 1024x4096 coefficients at Ra $> 10^{9}$ in 2D.

@@

In 2D, there are often multiple steady state solutions (e.g., 2-roll and 3-roll
solutions) which have slightly different flow properties (heat transport, etc.).
Even though the initial perturbations are very small, they shape the convective
transient and thus determine the nature of the steady state convection, at least in
the laminar regime.  In order to ensure that our results are not biased by differences
in flow structure, we ran the simulations using distinct random temperature perturbations
so as to compare statistics in comparable flow fields.  In 3D, rolls are nonstationary over
convective timescales, and so these effects need not be considered there.

\section{Results}
While the differences in the fluxes in Fig. \ref{fig:time_trace} are small, it is important
to determine if the velocity fluctuations and point-by-point nonlinear transport are the same
in the evolved state.  Fig. \ref{fig:pdf_comparison} overlays the probability distribution functions
of the vertical and horizontal velocities, as well as the fully nonlinear portion of the convective
flux for the same case as is shown in Fig. \ref{fig:time_trace}.  The PDFs are quite similar visually,
and have a similarity of (XYZ) according to a Kolmogorov-Smirnov statistic.

In addition to getting the nonlinear dynamics mostly correct, we show that the BVP method retrieves the
proper temperature profile, see e.g., Fig. \ref{fig:temp_comparison}.  Here the BVP profile retrieves the
mean profile of the temperature to within 1\% accuracy, and temperature fluctuations in the two runs
have a similarity of (XYZ) according to a Kolmogorov-Smirnov statistic.

This method works across a broad range of supercriticality.  In Fig. \ref{fig:parameter_space_comparison},
we show measurements of the volume-averaged Nusselt number, Reynolds number, and temperature.
We use standard definitions of the Nusselt number and Reynolds numbers,
\begin{equation}
\text{Nu} = \frac{\angles{wT - (\text{Ra Pr})^{-1/2}\grad T}}{\angles{- (\text{Ra Pr})^{-1/2} \grad T}} =
1 + \frac{\angles{wT}}{-\Delta T}\sqrt{\text{Ra Pr}}, \qquad \text{Re} = \frac{|\bm{u}| L_z}{\nu},
\end{equation}
where $\Delta T = T(z = 1/2) - T(z = -1/2)$ is the evolved temperature difference
between the top and bottom plates.  This form of the Nusselt number is valid even when
the system is not yet in flux equilibrium, and reduces to the standard fixed flux definition
of Nu  = $[1 - \angles{wT} / P]^{-1}$ \cite{johnston&doering2009}. We find a scaling law of
Nu $\propto \text{Ra}^{2/3}$, much steeper than a standard 2/7 or 1/3 scaling law
\cite{johnston&doering2009}. Furthermore, we find that Re$\propto \text{Ra}^{0.425}$. The average temperature
approaches the value at the top of the domain as Ra increases.

The final morphology of the flows is very important in determining the exact value of Nu and Re.  In general,
a two-roll state vs. a three-roll state will have entirely different statistics -- different Nu, Re, and average
temprature (and as a result, different size boundary layers).  Thus, in 2D studies, it is essential to study flows
of a similar morphology in order to see a clear trend.

@@

>>>>

\begin{abstract}
We present a method for coupling Boundary value problems (BVPs) with Initial value problems (IVPs)
in order to achieve thermally converged convective solutions on dynamical timescales, rather than the
long thermal timescale.  We demonstrate the similarity between the solution reached by BVP and the
solution reached by a long thermal rundown of the IVP, and demonstrate that this method works at a
large range of supercriticalities.  We use this method to achieve converged solutions at high Ra,
and discuss its extension to more complex scenarios, such as stratified, compressible convection.
\end{abstract}
\maketitle

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales. Granules on the
solar surface overturn on the order of 10 minutes, whereas deep motions in the Sun are likely at
low Mach number and constrained by the solar rotation rate of $\sim$1 month.  
Both of these dynamical timescales are vastly shorter than the Sun's average timescale of energy transport,
which is the Kelvin-Helmholtz timescale of nearly $3 \times 10^7$ \emph{years} \cite{stix2003}. 
As simulations aim to model natural convection
by increasing into the high-Rayleigh Number (Ra) regime, where diffusive timescales are much
longer than dynamical timescales \cite{anders&brown2017}, achieving converged simulations will 
require runs which span a greater number of convective timescales in order to thermally converge.
Furthermore, with increasing Ra and decreasing diffusivities, motions become more turbulent
and require finer grid meshes and shorter timesteps to resolve turbulent motions, increasing the simulation
time required for each overturn timescale.  
These two effects conspire to make achieving thermally converged, high-Ra, astrophysically interesting
simulations an intractable problem using modern numerical tools.

Prior studies of stratified convection in which a convective layer lies between stable layers
have used the knowledge of Mixing Length Theory (MLT) to adjust the initial thermodynamic structure of
the atmosphere to a state which is closer to the adiabat chosen by convection \cite{brandenburg&all2005}.
However, many modern studies of convection do not contain stable layers above and below the convection
zone, and the presence of hard boundaries and the boundary layers that they form make it difficult to
know the correct evolved adiabat \emph{a priori}.

Here we present a method for using simple boundary value problems (BVPs), 
along with information about the evolved flow fields,
to fast-forward the slow thermal evolution of convecting simulations.  
We run two sets of experiments: one in which we allow convective simulations to evolve for a
full thermal timescale before taking measurements, and another in which we employ a fast,
BVP technique which occurs on dynamical timescales. We compare these two sets of simulations to
show the validity of the BVP technique.  We use the BVP technique to run simulations
at high Ra, in the regime where running for thermal timescales becomes computationally intractable.

\section{Experiment}
\label{sec:experiment}
We adopt the Oberbeck-Boussinesq approximation.  Under this choice, the
fluid has constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The variables of the fluid that are evolved are the velocity,
$\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$, the temperature $T = T_0 + T_1$, and the pressure, $P$.
The density of the fluid is a constant $\rho_0$, except on the
term where the constant gravitational acceleration, $\bm{g} = - g\hat{z}$ acts in the vertical momentum equation, 
in which case it is $\rho = \rho_0(1  - \alpha T_1)$.  
Under these choices, the equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:dim_incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
We non-dimensionalize these equations such that length scales are in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and time is in units of the freefall timescale ($L_z / v_{\text{ff}}$), where the freefall velocity is
$v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$.
We further re-arrange the momentum equation by introducing a reduced kinematic pressure,
$\varpi \equiv P / \rho_0 + \phi + |\bm{u}|^2 / 2$, where $\bm{g} = -\grad \phi$ is the gravitational potential.
Despite containing a nonlinear $\bm{u}$ component, the nature of $P$ in \RB convection, as a
Lagrange multiplier,
allows us to treat $\varpi$ as a linear unknown, without a need to resolve its different components.
We time evolve the equations in the form,
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega}
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity. We find Eq. (\ref{eqn:bouss_momentum}) to
be slightly faster numerically than the standard form in Eq. (\ref{eqn:dim_bouss_momentum}).
The dimensionless control parameters are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\chi} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\chi}, \qquad \text{Pr} = \frac{\nu}{\chi}.
\end{equation}
The dimensionless vertical extent of the domain is $z = [-1/2, 1/2]$, and at the boundaries
we impose no-slip, impenetrable boundary conditions such that $w = u = v = 0$ at $z = \pm 1/2$.
At the lower boundary, we employ a fixed flux condition such that $\partial T_1 / \partial z = 0$
at $z = -1/2$, and we impose a fixed temperature condition $T_1 = 0$ at $z = 1/2$. Both
horizontal directions are periodic, extending over a range $x, y = [0, \Gamma]$, where
the aspect ratio is $\Gamma = 2$. In our 2D cases, we only examine the $x$ and $z$ dimensions
and we set $v = \partial_y = 0$.

The chosen thermal boundary conditions at the upper and lower plates
determine key quantities of the evolved state.
Studies of incompressible, Boussinesq, \RB convection often
employ fixed temperature (Dirichlet) or fixed heat flux
(Neumann) boundary conditions at both plates.  
Dirichlet conditions represent plates of infinite conductivity,
whereas Neumann conditions model plates of finite conductivity.  
In both cases, choosing symmetric boundary conditions maintains overall system symmetry, 
and despite evolving towards different thermal structures, both types of conditions
transport heat in the same manner \cite{johnston&doering2009}.
Studies of convection which aim to model
astrophysical systems such as stars often employ a mixture of these
two types of boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017}.  
The flux at the lower boundary is fixed, modeling
the constant energy generation of the stellar core, 
while the outer boundary condition is held at a fixed temperature,
modeling the surface of a star which must output the energy generated internally.
This setup is a useful model for understanding natural
systems, but simulations which employ these boundary conditions suffer from a long 
thermal relaxation as the atmosphere loses energy and approaches the adiabat chosen by the
Dirichlet condition.  We choose these conditions in part to better understand them, and in
part because these conditions minimize the number of assumptions that must be made in
setting up the boundary value problem.  For this choice of boundary conditions, the
critical Ra is Ra$_{\text{crit}} = 1295.78$, and we will often speak in terms of the
supercriticality, $S = \text{Ra}/\text{Ra}_{\text{crit}}$.

@@

It is reasonable to assume that convection will evolve the mean temperature profile that
drives it over the course of a thermal diffusion time, $t_\chi = \mathcal{P}^{-1}$. Thus, as Ra
increases, the timescale for achieving a thermally converged mean temperature profile becomes intractable.
This prohibitively long thermal timescale in an Initial Value Problem (IVP)
can be skipped by coupling the IVP with a simple Boundary Value Problem
(BVP) solve. By using information about the convecting dynamical state of an unconverged solution,
a BVP can be used to solve for the evolved system state.
A comparison of a simulation running for the thermal timescale, and another where a BVP is used
to fast-forward atmospheric evolution, are shown in Fig \ref{fig:time_trace}a\&b.

The Boussinesq BVP in essence contains equations of hydrostatic balance and thermal equilibrium,
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi} - \angles{T_1}\hat{z} = \angles{\bm{u}\times\bm{\omega}},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\angles{wT_1} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1} = 0,
	\label{eqn:bouss_BVP_energy}
\end{gather}
where $\angles{A}$ represents a time- and horizontally averaged profile of the quantity $A$.  
These
equations arise from assuming that the atmosphere is in a steady state ($\partial_t \rightarrow 0$),
then taking time and horizontal averages of Eqns (\ref{eqn:bouss_momentum}\&\ref{eqn:bouss_energy}) and
neglecting terms that vanish due to the symmetry of the problem.
Convective flows
are perturbations around a thermal profile defined by these equations in the proper evolved, statistically stationary state.

@@

By definition, the profile of $F_{\text{conv}}$ is not in a time stationary state early in the IVP.  
In fact, as the atmosphere approaches the
isotherm specified by the upper (fixed $T$) boundary condition, the motions display an asymmetric flux as energy
leaks through the upper boundary condition (Fig. \ref{fig:time_trace}c) in order to reach a lower temperature state.
In order to construct the evolved convective flux from the current fluxes in the atmosphere,
we assume that the atmospheric dynamics have properly developed the thickness of the boundary layers.
In other words, if we know the quantities $F_{\text{conv}}$, $F_{\text{cond}} = \angles{-P \partial_z T}$,
and $F_{\text{tot}} = F_{\text{conv}} + F_{\text{cond}}$, we assume that the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}}
\label{eqn:bvp_ratios}
\end{equation}
early in the evolution are the same as they will be in the final steady state solution. By definition, the
evolved total flux through the atmosphere must be the same as the flux entering the atmosphere at the
bottom boundary, $F_{\text{tot, steady}} = F_{\text{bot}}$.  The proper convective flux profile to use in
Eqn. (\ref{eqn:bouss_BVP_energy}) is then 
$\angles{w T_1}_{\text{steady}} = F_{\text{tot, steady}} \cdot f_{\text{conv}}$.

In order to carry the converged amount of flux appropriately, 
the velocity field and temperature fluctuations in the system must be scaled by a quantity
$\xi(z) = \angles{w T_1}_{\text{steady}}/\angles{w T_1}$.  Thus, we multiply the velocities
and the thermal fluctuations, $T - \angles{T}$, by $\sqrt{\xi}$, such that the product of all fluctuations
(which carry the convective flux) are diminished by a factor of $\xi$.  This also means that
$\angles{\bm{u}\times\bm{\omega}}_{\text{steady}} = \xi \angles{\bm{u}\times\bm{\omega}}$. Once these profiles
are appropriately adjusted, they can be used in the BVP solve to find $\angles{T}$ and $\angles{\varpi}$.

In general, the BVP solve is completed in the following steps:
\begin{enumerate}
\item Run the convective IVP. Once the convection achieves a volume-averaged Re of $\sqrt{\text{Ra}/\text{Ra}_{\text{crit}}}$,
wait for 50 freefall time units.  
\item Start taking the averages of $F_{\text{conv}}$, $F_{\text{tot}}$, and $\angles{\bm{u} \times \bm{\omega}}$, waiting either
30 freefall time units or until the average profiles change by no more than 1 part in 1000 on a
given timestep, whichever is a more difficult criterion. This ensures that
the profiles being used in the BVP are steady and smooth, and that they sample the full behavior of the convection.
\item Construct $\angles{w T_1}_{\text{steady}}$, $\xi$, and $\angles{\bm{u}\times\bm{\omega}}_{\text{steady}}$
from the flux profiles.
\item Solve for $\angles{T_1}$ and \angles{\varpi} of the
evolved state.  Adjust the mean profiles in the IVP.
\item Multiply the velocity field and the fluctuations in $T_1$ about its horizontal average by $\sqrt{\xi}$ in the IVP. 
\item Continue running the IVP for freefall 50 time units to allow for the velocities to equilibrate to their new background state.
\item Repeat steps 2-6 to complete a second BVP and get closer to the right solution now that we
are far from the very unstable transient period.
\item Take averages for the amount of time specified for the given run in Appendix \ref{section:appendix_table}.
\end{enumerate}

\subsection{Numerics}
We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to time-evolve  
(\ref{eqn:incompressible})-(\ref{eqn:bouss_energy}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
The temperature field is decomposed as $T = T_0(z) + T_1(x, y, z, t)$
and the velocity is $\bm{u} = w\bm{\hat{z}} + u\bm{\hat{x}} + v\bm{\hat{y}}$.
In our 2D runs, $v = 0$.
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.

@@

\section{Results}
While Fig. \ref{fig:time_trace} shows us that the post-BVP and post-rundown solutions
exhibit similar system energies and fluxes, it is important to examine other aspects of
the evolved solution. The BVP mainly serves to adjust the thermodynamic state of the solution;
in \RB convection, this means that the evolved temperature profile is a good measure of whether
or not the BVP achieves the proper solution.

The temperature profiles of a 2D run at $S = 10^5$ 
for both the BVP solve and the long rundown are shown in
Fig. \ref{fig:temp_comparison}a.  While they are not perfectly aligned they are within less than
2\% of each other throughout the depth of the atmosphere, as shown in Fig. \ref{fig:temp_comparison}b.
While the mean temperature profile is interesting, the point-by-point fluctuations in the
temperature are what lead to convective transport, and thus examining the spread of temperature
and the allowed values that it can take are important to characterizing the system.  As seen
in Fig. \ref{fig:temp_comparison}c, aside from the noticeable difference in the mode of the
temperatures, the two runs exhibit largely the same spread of temperatures, with the rundown
achieving higher temperatures at the base of the atmosphere.  These two profiles do not pass
a Kolmogorov-Smirnov test for similarity (AUTHOR NOTE: we need to collaborate on different measures
of comparing PDFs).

In addition to examining the thermodynamics of the solutions, we are especially interested in
determining if the post-BVP solution appropriately captures key features of the nonlinear dynamics.
We show point-by-point frequency distributions of the horizontal velocity, vertical velocity,
and nonlinear transport in Fig. \ref{fig:pdf_comparison}.  All of these measures have a sharp
peak at zero due to the no-slip, impenetrable boundary conditions and the denser grid sampling
near the boundaries of our Chebyshev grid.  Regardless, we see that the velocities of the 
two solutions are largely consistent.  The nonlinear transport in Fig. \ref{fig:pdf_comparison}
shows that the post-BVP solution experiences more extreme events which carry a large amount of
flux than the rundown solution, which is to be expected since it is not quite in equilibrium,
as shown in Fig. \ref{fig:time_trace}d.  None of these profiles pass a KS test for similarity
(AUTHOR NOTE: see previous paragraph comment).

@@

The goodness of this method has been tested across a broad range of supercriticality.  We have
examined both Rundown and BVP solutions in the range $S = [10^{1/3}, 10^5]$ in 2D and
$S = [10^1, 10^4]$ in 3D (3D points to come).  We have also further explored up to $S = 10^6$
(AUTHOR NOTE: should we go up to $10^8$?  $10^7$ is easily in reach. Also, 3D data points are
sparse and will change -- need to rerun with proper BVP procedure).  We compare scalar, 
volume-averaged measurements of all of these runs in Fig. \ref{fig:parameter_space_comparison}.

In Fig. \ref{fig:parameter_space_comparison}a, we compare the Nusselt number scaling in 2D,
3D, and for BVP and rundown solutions.  We use a classic definition of the Nusselt number
to measure the heat transport in these systems, where
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - (\text{Ra Pr})^{-1/2}\grad T}}{\angles{- (\text{Ra Pr})^{-1/2} \grad T}}.
\end{equation}
At small values of $S$, we achieve a steady state solution with a clear value of Nu.  Near
$S \sim 10^3$, our simulations begin to achieve a horizontally oscillatory state.  Due to the
use of no-slip boundaries, these states never give way to full scale shearing states, but they
do reduce the overall heat transport through the system and cause the systems to oscillate between
periods of high- and low- heat transport.  The vertical lines represent the standard deviation 
of Nu associated with these oscillations, while the error bars show the drift of the mean
value of Nu over the averaging window.  We find that $\text{Nu} \propto \text{Ra}^{1/5}$,
a weaker scaling law than that seen in traditional \RB convection \cite{johnston&doering2009},
but similar to that seen in shearing states [cite Goluskin].
(AUTHOR NOTE: Ben, I think this is entirely because of the oscillatory states.  I wonder what
would happen if we included viscous fluxes...)

In Fig. \ref{fig:parameter_space_comparison}b, we measure the rms Reynolds number, where
$\text{Re} = \angles{|\bm{u}|} / \mathcal{R}$.  We find that this scales roughly as
$\text{Re} \propto \text{Ra}^{0.45}$, and find good agreement between BVP and rundown solutions.

In Fig. \ref{fig:parameter_space_comparison}c, we measure the volume averaged 
temperature of our solution, then subtract out the value at the upper (fixed $T$) boundary.
If the atmosphere were perfectly isothermal, this value would be exactly zero.  Thus, this is
a measure of the temperature jump across the boundary layers, and it provides insight both
into whether or not the temperature profiles of two runs are the same and into whether
the fluxes through the two systems are the same (it is in some ways an alternate measure of Nu
[cite King 2009 supp materials]).
We find that this measure is $\propto \text{Ra}^{-1/5}$, approaching zero, as expected.
